---
title: "Data 608 HW 4 LQ"
author: "Dmitriy Burtsev"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library
```{r}
library(pacman)
p_load(tidyverse)
p_load(dlookr)
library(DHARMa)
p_load(caret)
```

## Import and Data Cleaning

```{r}
# read in the training data
train = read_csv('https://raw.githubusercontent.com/MsQCompSci/Data621Group4/main/HW4/insurance_training_data.csv')
# read in the test data 
test = read_csv('https://raw.githubusercontent.com/MsQCompSci/Data621Group4/main/HW4/insurance-evaluation-data.csv')
```

## explore the training data
```{r}
train %>% glimpse()
train %>% describe()
```
## Some issues to address:
 remove the index column
 make INCOME, HOME_VAL, BLUEBOOK, and OLDCLAIM numeric
 remove '_z' from MSTATUS, SEX, EDUCATION, JOB
```{r}
# isolate the character columns in a variable
character_cols = train %>% select(where(is.character)) %>% colnames() 
```
 
## Train data
```{r}
# remove unwanted characters from character columns
x = train %>% mutate_at(.vars = character_cols
                   , .funs = gsub
                   , pattern = "[$|_z|,|<|z_]+"
                   , replacement = '') %>%

  # change home_val, bluebook, and oldclaim into numeric variables
  mutate_at(.vars = c('HOME_VAL', 'BLUEBOOK', 'OLDCLAIM', 'INCOME')
            , .funs = as.numeric) %>% 
  
  # change character strings to lowercase
  mutate_if(is.character, tolower) %>%
  
  # change characters to factors
  mutate_if(is.character, as.factor) %>%
  
  # mutate urban/rural
  mutate(URBANICITY = if_else(URBANICITY == 'highly urban/ urban'
                             , 'urban'
                             , 'rural')) %>%
  # drop the INDEX column
  select(-INDEX) %>% 
  
  # drop nas
  drop_na() 

# isolate the character columns in a variable (test)
character_cols_test = test %>% select(where(is.character)) %>% colnames()
# determine NA columns
na_cols_test = test %>% describe() %>% filter(na > 0) %>% select(variable) %>% pull()
```

## Test data
```{r}
# remove unwanted characters from character columns
y = test %>% mutate_at(.vars = character_cols_test
                        , .funs = gsub
                        , pattern = "[$|_z|,|<|z_]+"
                        , replacement = '') %>%
  
  # change home_val, bluebook, and oldclaim into numeric variables
  mutate_at(.vars = c('HOME_VAL', 'BLUEBOOK', 'OLDCLAIM', 'INCOME')
            , .funs = as.numeric) %>% 
  
  # change character strings to lowercase
  mutate_if(is.character, tolower) %>%
  
  # change characters to factors
  mutate_if(is.character, as.factor) %>%
  
  # mutate urban/rural
  mutate(URBANICITY = if_else(URBANICITY == 'highly urban/ urban'
                              , 'urban'
                              , 'rural')) %>%
  # drop the INDEX column
  select(-INDEX) %>%
  
  # drop NAs in the columns other than response columns
  drop_na(all_of(na_cols_test))
```

## Check dataframe after cleaning
```{r}
# check the df after cleaning
x %>% glimpse()

# check to make sure binary variables only have two unique values
x %>% diagnose() %>% 
  select(variables, unique_count) %>% 
  filter(unique_count == 2)

# check which columns contain NA (should be 0)
na_columns = x %>% diagnose() %>% 
  filter(missing_count > 0) %>% 
  select(variables) %>% pull()
```

## Exploratory Analysis
## Distribution of Response Variables
```{r}
# plot the distribution  of TARGET_AMT in train
train %>% drop_na() %>% ggplot() + aes(TARGET_AMT) + 
  geom_histogram(bins = 50, color = 'black', fill = 'darkgray') +
  theme_light() + theme(panel.border = element_blank())

# look at the distribution of the TARGET_FLAG binary variable
train %>% group_by(TARGET_FLAG) %>% summarise(n = n()) %>% ggplot() +
  aes(TARGET_FLAG %>% as.factor(), n) + geom_col(color = 'black'
                                                 , fill = 'darkgray') +
  theme_light() + theme(panel.border = element_blank())
```

## Outliers
```{r}
# search for any outliers in the data
x %>% diagnose_outlier()

# identify variables w/ outliers ratio > 5%
x %>% plot_outlier(diagnose_outlier(x) %>%
                     filter(outliers_ratio >= 0.5) %>%
                     select(variables) %>%
                     unlist()) 

# BLUEBOOK variable distribution 
x %>% ggplot() + aes(BLUEBOOK) + 
  geom_histogram(color = 'black', bins = 50)

# BLUEBOOK variable distribution normalized
x %>% ggplot() + aes(BLUEBOOK %>% log()) + 
  geom_histogram(color = 'black', bins = 50)
```

## Correlation
```{r}
# filter correlations in the variables for under/over -0.4 and 0.4
x %>% correlate() %>% filter(coef_corr <= -0.4|
                               coef_corr >= 0.4)

# plot the variables with highest correlations
x %>% plot_correlate(TARGET_AMT, TARGET_FLAG
                     ,HOMEKIDS, KIDSDRIV
                     , AGE, INCOME
                     , OLDCLAIM, CLM_FREQ)
```

## Linear Model
```{r}
# the first linear model will regress the TARGET_AMT 
# on all variables in the data frame (minus the INDEX)
lm1 = lm(TARGET_AMT ~ ., data = x %>% select(-TARGET_FLAG))
lm1 %>% summary()
```

## Logistic Model
```{r}
glmOut <- glm(formula = TARGET_FLAG ~ . 
              , family = binomial(link="logit")
              , data = x %>% 
                select(-c(TARGET_AMT)))
glmOut

simulateResiduals(fittedModel = glmOut, n = 250) %>% plot()
```

